The shape of dark matter halos in the AURIGA simulations

### 2019-03-05 Paper/paper.tex
edit

%\begin{multicols}{2}
\begin{table*}
%\setlength{\tabcolsep}{3pt}
\begin{tabular}{|l|cc|cc|c|c|p{4cm}|}\hline
Reference&$q_{\rho}$&$s_{\rho}$&$q_{\phi}$&$s_{\phi}$&$R$&$\theta$&comment\\ \hline \hline
\citet{Olling_and_Merrifield_2000}& $\mathbf{1.00}$ & $\mathbf{0.80}$ & & & $\simeq 8$kpc & $0^{\circ}$&Method: Stellar dynamics and HI density. \\\hline
\citet{Law_and_Majewski_2009}&&&$\mathbf{0.83}$&$\mathbf{0.67}$& $\lesssim 60$kpc&$90^{\circ}$&Mid-axis orientation. Method: Sagittarius stream\\\hline
%
\citet{Law_and_Majewski_2010}&&&$\mathbf{0.99}$&$\mathbf{0.72}$& $[20$kpc,$60$kpc$]$&$90^{\circ}$&Mid-axis orientation, Method: Sagittarius stream\\\hline
%
\citet{Loebman_et_al._2012}&$\mathbf{1.00}$&$\mathbf{0.47}$&&&$\sim 20$kpc &$0^{\circ}$&Method: SDSS statistics\\\hline
%
\citet{Deg_and_Widrow_2013}&$0.72$&$0.28$&$0.82$&$0.40$&$[20$kpc,$60$kpc$]$&$90^{\circ}$& Mid-axis orientation. Method: Sagittarius stream\\\hline
%
\multirow{2}{*}{\citet{Vera-Ciro_and_Helmi_2013}}&&&$\mathbf{1.00}$&$\mathbf{0.90}$&$\lesssim 10$kpc&$0^{\circ}$ & Method: Sagittarius stream \& LMC \\
&&&$\mathbf{0.90}$&$\mathbf{0.80}$&$\gtrsim 10$kpc&$90^{\circ}$& Mid-axis orientation on the outside. \\\hline
%
\citet{Bovy_et_el._2016}&$\mathbf{0.95}$&$\mathbf{0.95}$&&&$\lesssim 20$kpc&$90^{\circ}$ & Method: Stellar streams\\\hline
%
\citet{Bowden_et_al._2016}&&&$\mathbf{[0.5,0.66]}$&$\mathbf{[0.5,0.66]}$&$[5$kpc,$10$kpc$]$&$90^{\circ}$& Weak constraint on prolate halo. Method: SDSS stars dynamics.\\\hline
%
\multirow{2}{*}{\citet{Banerjee_and_Chanda_2011}}&$\mathbf{1}$&$\mathbf{1}$&&&$9$kpc&$0^{\circ}$&Method: HI gas. \\
&$\mathbf{0.5}$&$\mathbf{0.5}$&&&$24$kpc&$0^{\circ}$&Monotonical change between radial regimes.\\\hline
%
\cite{Johnston_et_al._2005}&$\mathbf{1}$&$\mathbf{[0.83-0.92]}$&&&$\lesssim 60$kpc&$0^{\circ}$&Method: Sagittarius stream\\\hline\hline
\end{tabular}
\caption{(TODO: compute analogues in isopotential or isodensity according to Binney and Tremaine)}
\end{table*}
%\end{multicols}

@@

\subsection{The orientation of the principa axes}

One of the principal assumptions of observational models of the MW's
DM halo is that its minor axis is perfectly aligned with the disk
axis. Although this is a reasonable assumption to guarantee the
stability of the galactic disk in simplified models of isolated
galaxies, it may not be the case for galaxies evolved in the whole
cosmological context nor at every radii at which the shape is
sampled.

Therefore, it is of special interest to us to examin the strength of
this alignment assumption in the context of simulations. For this
purpose, we sampled the shape at 5 different radii and plotted the
axial directions, as well as the disk direction in the \textbf{name}
diagrams \cite{} (explain diagrams). Explain disk direction
calculation.

We found that there is not a representative example of what happens in
terms of alignments. We found that the majority of the disks are
aligned with the minor axis of their DM halo within $\approx 30^o$; in
some special cases, this alignment was almost perfect and in some
other cases, the DM halo minor axis changed substantially to be able
to determine an alignment. In figure \ref{fig:alignment} we present
these three cases. \textbf{occurrencies for each case?}

@@

\begin{figure}
  \centering
  \subfloat[Perfectly aligned Axes]{\includegraphics[width=1\columnwidth]{./pics/well_axes.png}}
  \hfill
  \subfloat[Somewhat aligned Axes]{\includegraphics[width=1\columnwidth]{./pics/rotating_axes.png}}
  \hfill
  \subfloat[Chaotic Axes]{\includegraphics[width=1\columnwidth]{./pics/chaotic_axes.png}}
  \hfill
  \caption{Star: Minor Axis
    Triangle: Medium Axis
    Square: Major Axis
    Color: Radii at which shape was sampled (show radii sampled)
    Contour color: If orientation is above or below in this projection
    Cross (black): Orientation of the stellar disk
    Conclusion: Axes (minor) are not (generally) aligned with the
    stellar disk nor are they usually aligned with each other from
    different radii. We show some cases that may happen.} 
  \label{fig:alignment}
\end{figure}

@@

\subsection{Comparison with observational constraints}
To be able to compare our results with observational values, we must relate the calculated quanties with their corresponding isopotential analogue, in which observational constraints are usually presented. For this purpose, we run a simple algorithm to find an approximation of the shape of the isopotential contour. Here, we calculate the mean and standard deviation of the potential over a spherical shell of width equals to $10\%$ of the radius at which it is sampled. Then, we calculate the inertia tensor of particles with potential within $1\sigma$ around the mean potential and calculate its triaxial characterization with the reduced inertia tensor. We repeat the process of calculating the potential mean and standard deviation until convergence is achieved with tolerance of $10^{-4}$. We repeat this process for the different radii from table \ref{tabe:isopotential}. \\

>>>>

@@

\subsection{DM halo - Stellar Disk Alignment}

A common assumption in observational models of the MW DM halo is that
its minor axis is perfectly aligned with the disk axis.
Although it is a reasonable assumption to guarantee the
stability of the galactic disk in simplified models of isolated
galaxies, this might not hold in an explicit cosmological context.

To examine the validity of this assumption  we sampled the shape at 5
different radii and plotted  directions in a cartesiian coordinate
system where the $z$-axis always corresponds to the minor axis
measured at the virial radius.

{\bf en Este caso hace falta la figura con la distribuci\'on integrada de cos theta}.
Figure \ref{fig:alignment} is the cumulative distribution of $\cos\theta$, where $\theta$ is the 
summarizes our results. 
We find that the majority of the disks are aligned with the minor axis of their DM halo within $\approx 30^{\circ}$.
in some special cases, this alignment was almost perfect and in some
other cases, the DM halo minor axis changed substantially to be able
to determine an alignment. In figure \ref{fig:alignment} we present
these three cases. \textbf{occurrencies for each case?}

%---AVANCE

@@

\begin{figure*}
  \centering
  \subfloat[Perfectly aligned Axes]{\includegraphics[width=0.6\columnwidth]{./pics/well_axes.png}}
  \hfill
  \subfloat[Somewhat aligned Axes]{\includegraphics[width=0.6\columnwidth]{./pics/rotating_axes.png}}
  \hfill
  \subfloat[Chaotic Axes]{\includegraphics[width=0.6\columnwidth]{./pics/chaotic_axes.png}}
  \hfill
  \caption{Star: Minor Axis
    Triangle: Medium Axis
    Square: Major Axis
    Color: Radii at which shape was sampled (show radii sampled)
    Contour color: If orientation is above or below in this projection
    Cross (black): Orientation of the stellar disk
    Conclusion: Axes (minor) are not (generally) aligned with the
    stellar disk nor are they usually aligned with each other from
    different radii. We show some cases that may happen.} 
  \label{fig:alignment}
\end{figure*}

@@

\subsection{Comparison with observational constraints}

\begin{table*}
\begin{tabular}{|l|cc|cc|c|c|p{4cm}|}\hline
Reference&$q_{\rho}$&$s_{\rho}$&$q_{\phi}$&$s_{\phi}$&$R$&$\theta$&comment\\ \hline \hline
\citet{Olling_and_Merrifield_2000}& $\mathbf{1.00}$ & $\mathbf{0.80}$ & & & $\simeq 8$kpc & $0^{\circ}$&Method: Stellar dynamics and HI density. \\\hline
\citet{Law_and_Majewski_2009}&&&$\mathbf{0.83}$&$\mathbf{0.67}$& $\lesssim 60$kpc&$90^{\circ}$&Mid-axis orientation. Method: Sagittarius stream\\\hline
%
\citet{Law_and_Majewski_2010}&&&$\mathbf{0.99}$&$\mathbf{0.72}$& $[20$kpc,$60$kpc$]$&$90^{\circ}$&Mid-axis orientation, Method: Sagittarius stream\\\hline
%
\citet{Loebman_et_al._2012}&$\mathbf{1.00}$&$\mathbf{0.47}$&&&$\sim 20$kpc &$0^{\circ}$&Method: SDSS statistics\\\hline
%
\citet{Deg_and_Widrow_2013}&$0.72$&$0.28$&$0.82$&$0.40$&$[20$kpc,$60$kpc$]$&$90^{\circ}$& Mid-axis orientation. Method: Sagittarius stream\\\hline
%
\multirow{2}{*}{\citet{Vera-Ciro_and_Helmi_2013}}&&&$\mathbf{1.00}$&$\mathbf{0.90}$&$\lesssim 10$kpc&$0^{\circ}$ & Method: Sagittarius stream \& LMC \\
&&&$\mathbf{0.90}$&$\mathbf{0.80}$&$\gtrsim 10$kpc&$90^{\circ}$& Mid-axis orientation on the outside. \\\hline
%
\citet{Bovy_et_el._2016}&$\mathbf{0.95}$&$\mathbf{0.95}$&&&$\lesssim 20$kpc&$90^{\circ}$ & Method: Stellar streams\\\hline
%
\citet{Bowden_et_al._2016}&&&$\mathbf{[0.5,0.66]}$&$\mathbf{[0.5,0.66]}$&$[5$kpc,$10$kpc$]$&$90^{\circ}$& Weak constraint on prolate halo. Method: SDSS stars dynamics.\\\hline
%
\multirow{2}{*}{\citet{Banerjee_and_Chanda_2011}}&$\mathbf{1}$&$\mathbf{1}$&&&$9$kpc&$0^{\circ}$&Method: HI gas. \\
&$\mathbf{0.5}$&$\mathbf{0.5}$&&&$24$kpc&$0^{\circ}$&Monotonical change between radial regimes.\\\hline
%
\cite{Johnston_et_al._2005}&$\mathbf{1}$&$\mathbf{[0.83-0.92]}$&&&$\lesssim 60$kpc&$0^{\circ}$&Method: Sagittarius stream\\\hline\hline
\end{tabular}
\caption{(TODO: compute analogues in isopotential or isodensity according to Binney and Tremaine)}
\end{table*}
%\end{multicols}

To be able to compare our results with observational values, we must relate the calculated quanties with their corresponding isopotential analogue, in which observational constraints are usually presented. For this purpose, we run a simple algorithm to find an approximation of the shape of the isopotential contour. Here, we calculate the mean and standard deviation of the potential over a spherical shell of width equals to $10\%$ of the radius at which it is sampled. Then, we calculate the inertia tensor of particles with potential within $1\sigma$ around the mean potential and calculate its triaxial characterization with the reduced inertia tensor. We repeat the process of calculating the potential mean and standard deviation until convergence is achieved with tolerance of $10^{-4}$. We repeat this process for the different radii from table \ref{tabe:isopotential}. \\
